---
title: "Stoma_Project"
author: "Valentina Stoma"
date: "11/26/2019"
output: pdf_document
---




```{r}

library(rvest)
library(stringr)
library(tidyverse)
library(tidyr)
library(lubridate)
library(RSQLite)
library(sqldf)
library(readxl)
library(ggpubr)
library(usmap)
library(ggplot2)

```



```{r}
#reading in the files 
parks<-read_csv("Wiki_parks.csv")
#deleting unneccessary columns that are an accessaory of using Import.io
parks<- parks[,-c(1,3)]
head(parks)
```




```{r}
#creating a separate column to contain the copy of the  Area column for further modifications 
# and changing the format of the Date_estblished column into Date 
parks <- parks %>%
  mutate(Area_km = Area) %>%
  mutate(Date_established = mdy(Date_established))

head(parks)
#cleaning up the area column in order to obtain only the acres in one column 
clean_parks<- parks %>%
  mutate(Area = str_replace(Area, "acres \\(", " ")) %>%
  mutate(Area = str_replace(Area, "km", "")) %>%
  separate(., Area, into=c("Area", "Area_new"), sep=" ") 
#getting rid of unnecessary columns 
clean_parks<-clean_parks[,-c(5,6)]
#My dplyr package is conflicting with some other packcages so for some of the 
#functions I need to specify to which package it belongs. 
clean_parks<- clean_parks %>%
  dplyr::rename(Area_acres = Area)
```

Issue arose with date format for the introduction into the SQL - needed to change it into charachter in the following manner in order to insert into SQL 
```{r}
#this format change allows me to have normal character format that will
#be read correctly in the SQL database
clean_parks$Date_established<- as.character(as.Date(as.character(clean_parks$Date_established), 
                                                    format = "%Y-%m-%d"))

#Using this str_trim, I trim out any lagging and leading spaces, which prohibit recognition
#of the same names in different table in the database
#additionally, deleting any spaces in the acres size for proper formatting. 
clean_parks <- clean_parks %>%
  mutate_if(is.character, str_trim) %>%
  mutate(Area_acres = str_replace(Area_acres, ",", ""))
#moreover, I am rounding all the acres sizing of the national parks for universal fomatting. 
clean_parks<-clean_parks %>%
  mutate(Area_acres = str_replace_all(Area_acres, "[.*+]", ""))
```



# 10 MOST VISITED PARKS IN THE US 
```{r}
#reading in the website 
visitors<- read_html("https://www.nps.gov/aboutus/visitation-numbers.htm")

#extracting names, visitors from the read in website
names <-visitors %>%
  html_nodes(".CS_Element_Custom~ .CS_Element_Custom+ .CS_Element_Custom thead+ tbody td:nth-child(2)") %>%
  html_text

number <- visitors %>%
  html_nodes("#cs_control_5546313 td~ td+ td") %>%
  html_text()


#combining the extracted information into a dataframe. 
most_visits<- data.frame(names, number)
#clean up the creatde dataframe to have appropriate formats for number and characters, 
#as well as formatting the numbers to not have any commas. For the universal notation in this project, 
#the names of the National park do not contain any other identifications except its actual name. 
most_visits<-most_visits %>%
  mutate(number = as.character(number)) %>%
  mutate(number = str_replace_all(number, ",", "")) %>%
  mutate(number = as.numeric(number)) %>%
  mutate(names = str_replace(names, "National Park", "")) %>%
  dplyr::rename(visitor_number = number) %>%
  dplyr::rename(Park_name = names)
#Deleting any lagging and leading spaces. 
most_visits <- most_visits %>%
  mutate_if(is.character, str_trim)
head(most_visits)
```

# LIST OF YOSEMITE ACCIDENTS OVER THE YEARS 
```{r}
#reading in the web page with the accidents. 
yosemite_accident<-read_html("https://www.nps.gov/yose/blogs/psarblog.htm")

#extracting the necessary information from the appropriate nodes. 
Yosemite_accident_dates<-yosemite_accident %>%
  html_nodes(".date") %>%
  html_text()

description<-yosemite_accident %>%
  html_nodes(".slug p") %>%
  html_text()
#creating a list of just the names of the park these accidents refer to. 
yose_name<-replicate(69, "Yosemite")
#this dataframe will not be included in the database, this is an example 
#of missing data from other National Parks, and should be supplemented later on. 
yose_acc<- data.frame(yose_name, Yosemite_accident_dates, description)
#modyifying the format of the date column as Date. 
yose_acc<-yose_acc %>%
  mutate(Yosemite_accident_dates = mdy(Yosemite_accident_dates)) %>%
  dplyr::rename(accident_description = description)

#saving the dataframe to an appropriate name and modying the date format to be read into the 
#SQL database in case in the future it will be supplemented with more national park accident data. 
yosemite_accidents<- yose_acc
yosemite_accidents$Yosemite_accident_dates<- as.character(as.Date(as.character(yosemite_accidents$Yosemite_accident_dates), 
                                                                  format = "%Y-%m-%d"))
  
#this plot is generated to demostrate when the accidents occur and to show that 
#the data can be used in the analysis if more is collected accessibly by the national parks. 
ggplot(yose_acc, aes(Yosemite_accident_dates)) + 
  geom_histogram() + 
  ggtitle("Counts of Accident Reports over the years in Yosemite")

```

# YELLOWSTONE TRAIL NAME and DIFFICULTY INFORMATION 
```{r}
#reading the web site containing the trail information. 
yellowstone_trail<-read_html("https://www.yellowstonenationalparklodges.com/connect/yellowstone-hot-spot/trail-mix-12-awesome-day-hikes-in-yellowstone/")

#findign and reading in the nodes that have the inforamtion about names and trail difficulty. 
trail_names<-yellowstone_trail %>%
  html_nodes("h3 strong") %>%
  html_text()

yellowstone_diff<- yellowstone_trail %>%
  html_nodes("p:nth-child(18) , p:nth-child(23) , p:nth-child(34) , p:nth-child(39) , p:nth-child(44) , p:nth-child(49) , p:nth-child(54) , p:nth-child(60) , p:nth-child(64) , p:nth-child(29) , p:nth-child(7) , p:nth-child(13)") %>%
  html_text()


yellow_loc<- yellowstone_trail %>%
  html_nodes("p:nth-child(65) , p:nth-child(60) , p:nth-child(55) , p:nth-child(50) , p:nth-child(45) , p:nth-child(40) , p:nth-child(35) , p:nth-child(30) , p:nth-child(24) , p:nth-child(19) , p:nth-child(14) , p:nth-child(8)") %>%
  html_text()
#creating a list containig the name of the national park for which I obtained the trail 
#information. 12 is the number of trails I found from one source.
yellow_name<-replicate(12, "Yellowstone")
yellow_trails<- data.frame(yellow_name,trail_names, yellowstone_diff, yellow_loc)

#changing the content of one of the columns because of the way that the nodes read into R
yellow_trails <- yellow_trails %>%
  mutate(yellowstone_diff = str_replace(yellowstone_diff, "Location: Canyon Area", "Level of Difficulty: Strenuous"))
#cleaning up the obtained dataframe to reduce text in the level of difficulty column 
#and to standardize the classification of trail level. Additioanlly, changing the names 
#of the columns to accomodate addition of other national park trail information. 
yellowstone_trails <- yellow_trails %>%
  mutate(yellowstone_diff = str_replace(yellowstone_diff, "Level of [Dd]ifficulty:", "")) %>%
  mutate(yellowstone_diff = str_replace(yellowstone_diff, "Difficulty:", "")) %>%
  dplyr::rename(difficulty_level = yellowstone_diff) %>%
  mutate(trail_names = str_to_sentence(trail_names)) %>%
  mutate(yellow_loc = str_replace(yellow_loc, "Location: ", "")) %>%
  dplyr::rename(trail_location = yellow_loc) %>%
  dplyr::rename(park_name = yellow_name) %>%
  dplyr::rename(yellowstone_trail_name = trail_names)


```


# Elevation of parks 
```{r}
#this dataframe was obtained from Import.io
#there is some missing values in the Mountain range column in the dataframe so just in case replcae it with NA 
eleva_w_parks<- read.csv("Elevation_wiki_parks-(Crawl-Run)---2019-11-30T145634Z.csv", na.strings=c("","NA"))
head(eleva_w_parks)
#Formatting the type of columns in the dataframe. 
elevation<-eleva_w_parks %>%
  mutate(Mim_elevation = as.character(Mim_elevation)) %>%
  mutate(Vertical.relief = as.character(Vertical.relief)) 
elevation= as.tibble(elevation) 

#renaming the column because fof the typo from Import.io 
elevation<-elevation %>%
  dplyr::rename(Min_elevation = Mim_elevation)

head(elevation)
```

We can calculate the final elevation at the peak level from the provided min_elevation (ground level for the park) and the vertical.relief (vertical chanhe experienced by the park). 
```{r}
#cleaning up the data further to only contain one column with clean values of
#min_elevation and vertical.relief  in feet
elev<- elevation %>%
  mutate(Min_elevation = str_replace_all(Min_elevation, "feet", " " )) %>%
  mutate(Min_elevation = str_replace(Min_elevation, "m", "")) %>%
  mutate(Min_elevation = str_replace(Min_elevation, "\\(", "")) %>%
  mutate(Min_elevation = str_replace(Min_elevation, "\\)", "")) %>%
  separate(., Min_elevation, into = c("Min_elevation_feet", "Min_elevation_m"), 
           sep = " ", remove = TRUE) 
elev<- elev[, -c(5)]

elev<-elev %>%
  mutate(Vertical.relief = str_replace_all(Vertical.relief, "feet", " " )) %>%
  mutate(Vertical.relief = str_replace(Vertical.relief, "m", "")) %>%
  mutate(Vertical.relief = str_replace(Vertical.relief, "\\(", "")) %>%
  mutate(Vertical.relief = str_replace(Vertical.relief, "\\)", "")) %>%
  separate(., Vertical.relief, into= c("Vertical_relief_feet", "Vertical.relief.m"), 
           sep = " ", remove = TRUE)
  
  
elevation_all_parks<- elev[, -c(6)]
head(elevation_all_parks)
```

Calculating the final elevation : 
```{r}
#finally, calcualting the final elevation level for the peaks from the presented data. 
elev_all_parks<-elevation_all_parks %>%
  mutate(Min_elevation_feet = str_replace_all(Min_elevation_feet, ",", "")) %>%
  mutate(Vertical_relief_feet = str_replace_all(Vertical_relief_feet, ",", "")) %>% 
 
  mutate(Min_elevation_feet = as.integer(Min_elevation_feet)) %>%
  mutate(Vertical_relief_feet = as.integer(Vertical_relief_feet)) %>%
  mutate(Peak_height = Min_elevation_feet + Vertical_relief_feet)
#final clean up of the datafram to contain the correct formatting, as well as
#trim away any leading and lagging spaces. 
elev_all_parks<- elev_all_parks %>%
  mutate(Park_name = as.character(Park_name)) %>%
  mutate(Peak_name = as.character(Peak_name)) %>%
  mutate(Mountain_range = as.character(Mountain_range)) %>%
  mutate_if(is.character, str_trim)
```



Plot, which demostrates the peak height for the provided national parks
```{r}
#plotting the range of peaks for all the national parks. Changing the size of the 
#text for the ticks in order for them to not overlap. 
peaks_plot<-ggbarplot(elev_all_parks, x = "Park_name", y = "Peak_height") + 
  coord_flip()
peaks_plot + theme(axis.text.y =element_text(size = 5))
```


# YOSEMITE TRAIL INFORMATION  
```{r}
#additional trail information from a different national park - Yosemite. 
#The website is read with rvest. 
yose_trails<-read_html("https://www.nps.gov/yose/planyourvisit/valleyhikes.htm")

yose_trail_names<- yose_trails %>%
  html_nodes("#cs_idCell2x1x1 a") %>%
  html_text()

diff_yose_trail<- yose_trails %>%
  html_nodes("tr+ tr td:nth-child(2)") %>%
  html_text()

#similar to the previous 
yose_name<-replicate(10, "Yosemite")
yosemite_trails<- data.frame(yose_name, yose_trail_names, diff_yose_trail)



```

Connecting the three dataframes to contain the trail and difficulty information. 
```{r}
#cleaning up the final dataframes in order to connect them into one trail information containing dataframe 
yellow_trail_connect<-yellowstone_trails[-4]
yellow_trail_connect <- as.tibble(yellow_trail_connect)

#making sure that the names are the same as in all other tibbles to connect into one dataframe
yellow_trail_connect <- yellow_trail_connect %>%
  dplyr::rename(park = park_name) %>%
  dplyr::rename(trail_name = yellowstone_trail_name) %>%
  dplyr::rename(difficulty = difficulty_level) 



yosemite_trails_connect<- yosemite_trails %>%
  dplyr::rename(park = yose_name) %>%
  dplyr::rename(trail_name = yose_trail_names) %>%
  dplyr::rename(difficulty = diff_yose_trail )

```


```{r}
#connecting the dataframes
trail_level<-rbind(yellow_trail_connect, yosemite_trails_connect)
head(trail_level)
```

```{r}
#the third data subset to add to the final trail iformation containing dataframe
olympic_trails<-read_csv("nationalparked-olympic-data-(Crawl-Run)---2019-12-01T191803Z.csv")
olympic_trails<- olympic_trails[,-c(1)]
olymp_names<-replicate(9, "Olympic")
olympic_trails_connect<-data.frame(olymp_names, olympic_trails) 
olympic_trails_connect <- olympic_trails_connect %>%
  dplyr::rename(park = olymp_names)
head(olympic_trails_connect)
```


```{r}
#adding the third data subset to the previosuly made up dataframe with trail information 
trail_level_parks<- rbind(trail_level, olympic_trails_connect)
#figuring out which difficulty levels need to be fomratted 
trail_level_parks %>%
  group_by(difficulty) %>%
  count()
#formatting the difficulty level to contain universal names 
trail_level_3<-trail_level_parks %>%
  mutate(difficulty = str_replace_all(difficulty, " ", "")) %>%
  mutate(difficulty = str_replace_all(difficulty, "ModeratelyEasy", "EasytoModerate")) %>%
  mutate(difficulty = str_replace_all(difficulty, "VeryEasy", "Easy")) %>%
  mutate(difficulty = str_replace_all(difficulty, "EasytoModerate", "Moderate")) %>%
  mutate(difficulty = str_replace_all(difficulty, "ModeratetoStrenuous", "Strenuous"))
#doubel checking that there are only desired level of difficulty. 
trail_level_3 %>%
  group_by(difficulty) %>%
  count()
trail_level_3<- trail_level_3 %>%
  mutate_if(is.character, str_trim)
head(trail_level_3)
```



# Oldest parks: 
can compare with the previous publicly available data. 
```{r}
#Reading in the web page with the 10 oldest parks in order to compare 
#it to the wikipedia provided information 
oldest<-read_html("https://www.worldatlas.com/articles/oldest-national-parks-in-the-united-states.html")

#extracting the nodes with the necessary information. 
oldest_names<- oldest %>%
  html_nodes("td:nth-child(2)") %>%
  html_text()

date_oldest<- oldest %>%
  html_nodes("td~ td+ td") %>%
  html_text()
#uniting as a dataframe.
old_parks<- data.frame(oldest_names, date_oldest)
#cleaning up the dataframe to separate the park and the state information into different columns. 
old_parks<-old_parks %>%
  separate(., oldest_names, into = c("Park","State"), sep = ",") 
#cleaning up the dates column in order to be able to format it into the date format. 
old_park_2<- old_parks %>%
  mutate(State = str_replace(State,"Wyoming-Montana-", "")) %>%
  mutate(date_oldest = str_replace_all(date_oldest, "st|th|nd", "")) %>%
  mutate(date_oldest = mdy(date_oldest))
  
#performing the same format change in order to upload it into the SQL database. 
old_park_2$date_oldest<- as.character(as.Date(as.character(old_park_2$date_oldest), format = "%Y-%m-%d"))
#trimming out the leading and lagging spaces. 
old_park<-old_park_2 %>%
  mutate_if(is.character, str_trim)
head(old_park_2)

```



# RECREATIONAL VISITS FOR 10 YEARS 
```{r}
#reading in the obtained annual visitations report. 
annual_report<-read_excel("Annual_park_1979_now.xlsx", col_names = TRUE, skip = 6 )
```

```{r}
#cleaning up the unnecesasry artificats of reading in the excel file into R
annual_report<-annual_report[-1, -c(11,12)]
#performing this name change for now to keep easy track of changed instead of numbers. 
colnames(annual_report)<- c("Park_name", "one", "two", "three", "four", "five", "six", "sev", "eight", "nine", "tin", "Average")

```


```{r}
#changing the layout and the structure of the dataframe to assign different year
#values to the rows instead of the seprate columns 
annual_report<- gather(annual_report, Year, Visitors, one:tin)
#performign the name change back tot he years
annual_report <- annual_report %>%
  mutate(Year = str_replace(Year, "one", "2009")) %>%
  mutate(Year = str_replace(Year, "two", "2010"))  %>%
  mutate(Year = str_replace(Year, "three", "2011"))  %>%
  mutate(Year = str_replace(Year, "four", "2012"))  %>%
  mutate(Year = str_replace(Year, "five", "2013"))  %>%
  mutate(Year = str_replace(Year, "six", "2014"))  %>%
  mutate(Year = str_replace(Year, "sev", "2015"))  %>%
  mutate(Year = str_replace(Year, "eight", "2016"))  %>%
  mutate(Year = str_replace(Year, "nine", "2017"))  %>%
  mutate(Year = str_replace(Year, "tin", "2018")) 

#formatting the values in the Average column to read as integer and round it as 
#some of the values have too many significant values 
annual_report<- annual_report %>% 
  mutate(Average = as.integer(Average)) %>%
  mutate(Average = format(round(Average, 2), nsmall = 2))

```



```{r}
#formatting the types of columns
annual_report$Year <- as.numeric(annual_report$Year)
annual_report$Average <- as.numeric(annual_report$Average)
head(annual_report)

```


```{r}
#modifying the internal formatting of the names of the parks to make sure 
#that the names are recognizable in different table in the SQL database. 
annual_report <- annual_report %>%
  mutate(Park_name = str_replace_all(Park_name, "NHP|NM|NP|NHS|NRA|EM|NS|MEM|RES|NNRA|NSR|NS|NRES|NB|& PRES|EM| & HS|NB|PKWY|NL|NRRA|NRR|W&SR|Park| R |EHP|NR", "" )) %>%
  mutate(Park_name = str_replace_all(Park_name, "Shiloh P", "Shiloh")) %>%
  mutate(Park_name = str_replace_all(Park_name, "Pea Ridge P", "Pea Ridge")) %>%
  mutate(Park_name = str_replace_all(Park_name, "Ozark R", "Ozark")) %>%
  mutate(Park_name = str_replace_all(Park_name, "Niobrara R", "Niobrara")) %>%
  mutate(Park_name = str_replace_all(Park_name,"Chickamauga & Chattanooga P", "Chickamauga & Chattanooga")) %>%
  mutate(Park_name = str_replace_all(Park_name, "Fredericksburg & Spotsylvania P", "Fredericksburg & Spotsylvania")) %>%
  mutate(Park_name = str_replace_all(Park_name, "Gettysburg P", "Gettysburg")) %>%
  mutate(Park_name = str_replace_all(Park_name,"Guilford Courthouse P", "Guilford Courthouse"))%>%
  mutate(Park_name = str_replace_all(Park_name,"Horseshoe Bend P", "Horseshoe Bend")) %>%
  mutate(Park_name = str_replace_all(Park_name,"Kennesaw Mountain P" ,"Kennesaw Mountain")) %>%
  mutate(Park_name = str_replace_all(Park_name,"Kings Mountain P", "Kings Mountain")) %>%
  mutate(Park_name = str_replace_all(Park_name,"Manassas P",  "Manassas"))%>%
  mutate(Park_name = str_replace_all(Park_name,"Richmond P", "Richmond" ))%>%
  mutate(Park_name = str_replace_all(Park_name,"River Raisin P", "River Raisin")) %>%
  mutate(Park_name = str_replace_all(Park_name,"Saint Croix R" , "Saint Croix"))%>%
  mutate(Park_name = str_replace_all(Park_name,"S&RR" , "")) %>%
  mutate(Park_name = str_replace_all(Park_name,"Vicksburg P", "Vicksburg"))
  
```


```{r}
#trim out all the unnecessary lagging and leading spaces that prevent 
#tables inteh database from joining on the same values of the park name. 
annual_report<-annual_report %>%
  mutate_if(is.character, str_trim)
head(annual_report)

```

Example of the kind of visualisation that can be used to address the chnages in the visitor number for different parks 
```{r}
#plot example of the analysis that can be done on visito number to certain parks for all the years. 
zio_plot<-annual_report %>%
  filter(Park_name == "Zion") %>% 
  ggplot() + geom_line(aes(x = Year, y = Visitors )) + 
  ggtitle("Zion visitor number over the years") 


 
```

```{r}
#subsetting the data arbitrary to only contain the to 100 most visited 
#parks irrespective of the park and year limitations - done for exampel visualizaztion purposes. 
plot100_most<-annual_report %>%
  group_by(Year) %>%
  arrange(desc(Visitors)) %>%
  head(100)

```
Demonstration of the most visited parks over the years, limited sample of data for this example plot.  
```{r}
#another plot as an example of the type of visualization that can be done on the changes in 
#visitors to the parks - in this case subsetting the data arbitrary at 100 most visited parks 
#all through out the ten year reported period, Becasue of this arbitrary cut off, some parks
#only have sporadic values, as they reached the significant level 
top100_plot<-ggplot(plot100_most, aes(x = Year, y = Visitors, color = Park_name)) + 
  geom_line() + 
  ggtitle("Visitor number over the years in top visited\n National Parks in the US ")
```



Initate the database creation 
```{r}
db <- dbConnect(SQLite(), dbname="NationalParkInfo.sqlite")
dbSendQuery(conn = db, "pragma foreign_keys=on;")
```



Clean parks
```{r}
#creating tables, identifying the primary key
# dbSendQuery(conn = db,"CREATE TABLE clean_parks (
#                        Name TEXT,
#                        State TEXT,
#                        Date_established DATE,
#                        Area_acres INTEGER,
#                        PRIMARY KEY (Name))
#                        WITHOUT ROWID")
```


```{r}
#inserting the data from the appropriate tables in to the 
# dbWriteTable(conn = db, name = "clean_parks", value = clean_parks, row.names=FALSE, append = TRUE)

```


```{r}
#example of the equery working on the provided data - 
clean_park_q<-dbGetQuery(db, "SELECT * FROM clean_parks")
head(clean_park_q)
```


Annual Report 
```{r}
# dbSendQuery(conn = db,"CREATE TABLE annual_report (
#                        Park_name TEXT,
#                        Average Integer,
#                        Year INTEGER,
#                        Visitors INTEGER,
#                        PRIMARY KEY (Park_name, Year))
#                        WITHOUT ROWID")
```

```{r}
# dbWriteTable(conn = db, name = "annual_report", value = annual_report, row.names=FALSE, append = TRUE)

```


Visist 
```{r}
# dbSendQuery(conn = db,"CREATE TABLE most_visits  (
#                        Park_name TEXT,
#                        visitor_number INTEGER,
#                        PRIMARY KEY (Park_name))
#                        WITHOUT ROWID")
```

```{r}
# dbWriteTable(conn = db, name = "most_visits", value = most_visits, row.names=FALSE, append = TRUE)

```


Elevation 
```{r}
  # dbSendQuery(conn = db,"CREATE TABLE elev_all_parks (
  #                      Park_name TEXT,
  #                      Peak_name TEXT,
  #                      Mountain_range TEXT,
  #                      Min_elevation_feet INTEGER,
  #                      Vertical_relief_feet INTEGER,
  #                      Peak_height INTEGER,
  #                      PRIMARY KEY (Park_name))
  #                      WITHOUT ROWID")
```

```{r}
# dbWriteTable(conn = db, name = "elev_all_parks", value = elev_all_parks, row.names=FALSE, append = TRUE)

```




Trail difficulty 
```{r}
# dbSendQuery(conn = db,"CREATE TABLE trail_level_3 (
#                      park TEXT,
#                      trail_name TEXT,
#                      difficulty TEXT,
#                      PRIMARY KEY (trail_name))
#                      WITHOUT ROWID")
```

```{r}
# dbWriteTable(conn = db, name = "trail_level_3", value = trail_level_3, row.names=FALSE, append = TRUE)

```


```{r}
# dbSendQuery(conn = db,"CREATE TABLE old_park_2 (
#                      Park TEXT,
#                      State TEXT,
#                      date_oldest DATE,
#                      PRIMARY KEY (Park))
#                      WITHOUT ROWID")
```

```{r}
#dbWriteTable(conn = db, name = "old_park_2", value = old_park_2, row.names=FALSE, append = TRUE)
```

```{r}
#visualization of the tables in the database. 
dbListTables(db)
```


```{r}
#This gets the visitors number for 2009 from the National Park Services 
#for only national parks in the US reported by the Wikipedia 
all_visits_2009<-dbGetQuery(db, "SELECT annual_report.Park_name, annual_report.Visitors, annual_report.Year
          FROM annual_report
          INNER JOIN clean_parks on clean_parks.Name = annual_report.Park_name
          WHERE annual_report.Year = 2009")
head(all_visits_2009)
```





```{r}
#this shows that the names are the same in the annual_report and clean_parks,
#and that SQL queries will work on these tables when they need to be joined. 
same_name<-dbGetQuery(db, "SELECT annual_report.Park_name FROM annual_report INTERSECT SELECT clean_parks.Name FROM clean_parks")
head(same_name)
```

```{r}
#This demonstrates that the names are identical in other dataframes as well. 
old_park_query<-dbGetQuery(db, "SELECT old_park_2.park FROM old_park_2 INTERSECT SELECT annual_report.Park_name FROM annual_report")
head(old_park_query)
```


```{r}
library(maps)
usa <- map_data("usa")

```



```{r}
#creating a separate column that has the abbreaviation of the states in order to be able to plot it with usmap.
clean_parks_abb<- clean_parks
clean_parks_abb$State_abb<-state.abb[match(clean_parks_abb$State,state.name)]

plot_us_clean<-clean_parks_abb %>%
  dplyr::rename(State_name = State)  %>%
  dplyr::rename(state = State_abb)
  
```


Connect with annual report with clean data info and plot the visitor population. 

```{r}
#I used this quaery as an example for future work that this database can be used for - 
#specifically, verification of the data obtained from open web soruces. T
#his cross reference verification is can be used with dates and other data. 
compare_dates<-dbGetQuery(db, "SELECT clean_parks.Name, clean_parks.State, clean_parks.Date_established,  old_park_2.date_oldest, old_park_2.Park
           FROM clean_parks 
            INNER JOIN old_park_2 on old_park_2.Park = clean_parks.Name ")
head(compare_dates)
```




```{r}
#combining the created two dataframes by the names of the park in order 
#to plot the average number of visitors
full_state_visit<-inner_join(clean_parks_abb, annual_report, by = c("Name" = "Park_name"))
```

```{r}
#This plot visualizes the average number of visitors per year to the national parks 
# in the corresponding states over the period of 10 years. 
full_state_visit<-full_state_visit %>%
  dplyr::rename(state = State_abb)
plot_usmap(data = full_state_visit, values = "Average", labels = TRUE)

```

